---
title: "Proof of Progress"
subtitle: "Data Expo 2018"
author: "Matt Higham, Erin Howard"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, cache=TRUE}
complete_df <- read.csv("~/Desktop/DataExpo2018/all_df_completesub.csv")

```

We are exploring how accurate precipation forecasts are for various cities in different seasons. We started our exploration by looking at two cities in Washington, but plan to expand our investigation nationally. Below are two plots designed to help visualize the accuracy of precipation forecasts in Hoquiam and Spokane, WA. The x-axis is the Probability of Precipitation (PoP), while the y-axis is the proportion of days in which rain was detected. The plots are faceted by the number of days between the date of the forecast and the date of measurement and are colored by season. We would expect perfect forecasts to lie on the black reference line. We are planning on making similar types of graphs for temperature forecasts.  

&nbsp;

&nbsp;

```{r, results = "hide", warning = FALSE, message = FALSE, cache=TRUE}
library(lubridate)
library(dplyr)
library(ggplot2)
names(complete_df)[8] <- "forecastValue"
precip <- subset(complete_df, weathermeas == "ProbPrecip")
head(precip)
precip$Date <- as.Date(precip$Date)
precip$DateofForecast <- as.Date(precip$DateofForecast)
precip$month <- month(as.POSIXlt(precip$Date))
Spok.precip <- subset(precip, city %in% c("Spokane") & month %in% c(3, 4, 5))
Spok.precip.summer <- subset(precip, city %in% c("Spokane") & month %in% c(6,7,8))
Spok.precip.fall <- subset(precip, city %in% c("Spokane") & month %in% c(9,10,11))
Spok.precip.winter <- subset(precip, city %in% c("Spokane") & month %in% c(12,1,2))
Spok.precip$ForecastTimeDay <- rep(c("Morning", "Evening"),
  length.out = nrow(Spok.precip)) 




head(Spok.precip)

Spok.precip$precipbinary <- as.numeric(Spok.precip$weatherval > 0.005)
Spok.precip.summer$precipbinary <- as.numeric(Spok.precip.summer$weatherval > 0.005)
Spok.precip.fall$precipbinary <- as.numeric(Spok.precip.fall$weatherval > 0.005)
Spok.precip.winter$precipbinary <- as.numeric(Spok.precip.winter$weatherval > 0.005)
Spok.precip$precipbinary


## for each forecastValue, want to figure out the proportion of days that it actually
## rained

Spok.precip$proprain <- (Spok.precip %>% group_by(forecastValue, 
  LengthForecastDayOnly, ForecastTimeDay) %>% 
    mutate(proprain = mean(precipbinary)))$proprain



Spok.precip$proprain_noMornEv <- (Spok.precip %>% group_by(forecastValue,
                                                           LengthForecastDayOnly) %>%
                                    mutate(proprain_noMornEv = mean(precipbinary)))$proprain_noMornEv

Spok.precip.summer$proprain_noMornEv <- (Spok.precip.summer %>% group_by(forecastValue,
                                                                         LengthForecastDayOnly) %>%
                                           mutate(proprain_noMornEv = mean(precipbinary)))$proprain_noMornEv

Spok.precip.fall$proprain_noMornEv <- (Spok.precip.fall %>% group_by(forecastValue,
                                                                         LengthForecastDayOnly) %>%
                                           mutate(proprain_noMornEv = mean(precipbinary)))$proprain_noMornEv

Spok.precip.winter$proprain_noMornEv <- (Spok.precip.winter %>% group_by(forecastValue,
                                                                     LengthForecastDayOnly) %>%
                                         mutate(proprain_noMornEv = mean(precipbinary)))$proprain_noMornEv


Spok.precip$Numerator_noMornEv <- (Spok.precip %>% group_by(forecastValue,
                                                            LengthForecastDayOnly) %>%
                                     mutate(Numerator_noMornEv = length(precipbinary)))$Numerator_noMornEv

Spok.precip.summer$Numerator_noMornEv <- (Spok.precip.summer %>% group_by(forecastValue, 
                                                                          LengthForecastDayOnly) %>%
                                            mutate(Numerator_noMornEv = length(precipbinary)))$Numerator_noMornEv

Spok.precip.fall$Numerator_noMornEv <- (Spok.precip.fall %>% group_by(forecastValue, 
                                                                          LengthForecastDayOnly) %>%
                                            mutate(Numerator_noMornEv = length(precipbinary)))$Numerator_noMornEv

Spok.precip.winter$Numerator_noMornEv <- (Spok.precip.winter %>% group_by(forecastValue, 
                                                                      LengthForecastDayOnly) %>%
                                          mutate(Numerator_noMornEv = length(precipbinary)))$Numerator_noMornEv


Spok.precip$season <- rep("Spring", nrow(Spok.precip))
Spok.precip.summer$season <- rep("Summer", nrow(Spok.precip.summer))
Spok.precip.fall$season <- rep("Fall", nrow(Spok.precip.fall))
Spok.precip.winter$season <- rep("Winter", nrow(Spok.precip.winter))

Spok.precip.all <- rbind(Spok.precip[,-c(20,22)], Spok.precip.summer, 
                         Spok.precip.fall, Spok.precip.winter)
                         

ggplot(data = Spok.precip.all, aes(x = forecastValue/100, y = proprain_noMornEv)) + 
  geom_point(aes(colour = season,alpha=Numerator_noMornEv)) + ylim(c(0, 1)) +
  facet_wrap(~LengthForecastDayOnly) +
  ggtitle("Spokane, WA") + 
  geom_smooth(aes(colour = season, weight = Numerator_noMornEv), method="lm", formula = y ~ x + I(x^2), 
              size = 0.5, se = FALSE) +
  #(aes(colour=season))
  scale_alpha_continuous(name = "Number of Obs") + 
  ylab("Proportion of Rainy Days") + 
  xlab("PoP") +
  geom_abline(intercept=0,slope=1)

Hoq.precip <- subset(precip, city %in% c("Hoquiam") & month %in% c(3, 4, 5))
Hoq.precip.summer <- subset(precip, city %in% c("Hoquiam") & month %in% c(6,7,8))
Hoq.precip.fall <- subset(precip, city %in% c("Hoquiam") & month %in% c(9,10,11))
Hoq.precip.winter <- subset(precip, city %in% c("Hoquiam") & month %in% c(12,1,2))
Hoq.precip$ForecastTimeDay <- rep(c("Morning", "Evening"),
                                   length.out = nrow(Hoq.precip)) 


subset(Hoq.precip, LengthForecastDayOnly %in% 4)$forecastValue

head(Hoq.precip)

Hoq.precip$precipbinary <- as.numeric(Hoq.precip$weatherval > 0.005)
Hoq.precip.summer$precipbinary <- as.numeric(Hoq.precip.summer$weatherval > 0.005)
Hoq.precip.fall$precipbinary <- as.numeric(Hoq.precip.fall$weatherval > 0.005)
Hoq.precip.winter$precipbinary <- as.numeric(Hoq.precip.winter$weatherval > 0.005)
Hoq.precip$precipbinary


## for each forecastValue, want to figure out the proportion of days that it actually
## rained

Hoq.precip$proprain <- (Hoq.precip %>% group_by(forecastValue, 
                                                  LengthForecastDayOnly, ForecastTimeDay) %>% 
                           mutate(proprain = mean(precipbinary)))$proprain



Hoq.precip$proprain_noMornEv <- (Hoq.precip %>% group_by(forecastValue,
                                                           LengthForecastDayOnly) %>%
                                    mutate(proprain_noMornEv = mean(precipbinary)))$proprain_noMornEv

Hoq.precip.summer$proprain_noMornEv <- (Hoq.precip.summer %>% group_by(forecastValue,
                                                                         LengthForecastDayOnly) %>%
                                           mutate(proprain_noMornEv = mean(precipbinary)))$proprain_noMornEv

Hoq.precip.fall$proprain_noMornEv <- (Hoq.precip.fall %>% group_by(forecastValue,
                                                                     LengthForecastDayOnly) %>%
                                         mutate(proprain_noMornEv = mean(precipbinary)))$proprain_noMornEv

Hoq.precip.winter$proprain_noMornEv <- (Hoq.precip.winter %>% group_by(forecastValue,
                                                                         LengthForecastDayOnly) %>%
                                           mutate(proprain_noMornEv = mean(precipbinary)))$proprain_noMornEv


Hoq.precip$Numerator_noMornEv <- (Hoq.precip %>% group_by(forecastValue,
                                                            LengthForecastDayOnly) %>%
                                     mutate(Numerator_noMornEv = length(precipbinary)))$Numerator_noMornEv

Hoq.precip.summer$Numerator_noMornEv <- (Hoq.precip.summer %>% group_by(forecastValue, 
                                                                          LengthForecastDayOnly) %>%
                                            mutate(Numerator_noMornEv = length(precipbinary)))$Numerator_noMornEv

Hoq.precip.fall$Numerator_noMornEv <- (Hoq.precip.fall %>% group_by(forecastValue, 
                                                                      LengthForecastDayOnly) %>%
                                          mutate(Numerator_noMornEv = length(precipbinary)))$Numerator_noMornEv

Hoq.precip.winter$Numerator_noMornEv <- (Hoq.precip.winter %>% group_by(forecastValue, 
                                                                          LengthForecastDayOnly) %>%
                                            mutate(Numerator_noMornEv = length(precipbinary)))$Numerator_noMornEv


Hoq.precip$season <- rep("Spring", nrow(Hoq.precip))
Hoq.precip.summer$season <- rep("Summer", nrow(Hoq.precip.summer))
Hoq.precip.fall$season <- rep("Fall", nrow(Hoq.precip.fall))
Hoq.precip.winter$season <- rep("Winter", nrow(Hoq.precip.winter))

Hoq.precip.all <- rbind(Hoq.precip[,-c(20,22)], Hoq.precip.summer, 
                         Hoq.precip.fall, Hoq.precip.winter)


ggplot(data = Hoq.precip.all, aes(x = forecastValue/100, y = proprain_noMornEv)) + 
  geom_point(aes(colour = season,alpha=Numerator_noMornEv)) + ylim(c(0, 1)) +
  facet_wrap(~LengthForecastDayOnly) +
  ggtitle("Hoquiam, WA") + 
  geom_smooth(aes(colour = season, weight = Numerator_noMornEv), method="lm", formula = y ~ x + I(x^2), 
              size = 0.5, se = FALSE) +
  #(aes(colour=season))
  scale_alpha_continuous(name = "Number of Obs") + 
  ylab("Proportion of Rainy Days") + 
  xlab("PoP") +
  geom_abline(intercept=0,slope=1)

```